---
title: "R21 Sterile Immunity Pseudobulk RNA-seq analysis"
output: html_document
date: "2023-05-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      root.dir = "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/Kalifabougou Sterile Immunity Manuscript/R21-sc-studies/DE_analysis_scrnaseq/")
```

## Objective

Perform single cell DGE between conditions for each cell type using pseudobulk method.

References:
https://hbctraining.github.io/scRNA-seq/lessons/pseudobulk_DESeq2_scrnaseq.html

https://www.nature.com/articles/s41467-021-25960-2/figures/1


# Load packages
```{r load packages}
library(Seurat) 
#library(SeuratWrappers)
library(patchwork)
library(SeuratDisk)
library(tidyverse)
library(magrittr)
library(googledrive)
library(ggpubr)
library(scales)
library(RColorBrewer)
library(viridis)
library(glue)
library(MAST)
library(Signac)
library(Matrix.utils)
library(apeglm)
#library(scater) #cannot install this package for some reason
library(DESeq2)
```

### Import refmapped Seurat object from googledrive

```{r import refmapped seurat from gdrve}
# Import here from Google Drive
# temp <- tempfile(fileext = ".rds")
# dl <- drive_download(
#   as_id("10EKJ-eE2_oouFiHTE5vwMfLF6qVZj_kc"), path = temp, overwrite = TRUE)
# refmapped_seurat <- readRDS(file = dl$local_path)
```

```{r import refmapped seurat from local path}
refmapped_seurat <- readRDS(file = "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/Kalifabougou Sterile Immunity Manuscript/R21-sc-studies/data/refmapped_and_dim_red_no_sampling_all15_minusK613_05172023.rds")
```


### subset and extract the cells from a Seurat object

```{r, subset and extract from seurat}
# Extract raw counts and metadata to create SingleCellExperiment object
counts <- refmapped_seurat@assays$RNA@counts 

metadata <- refmapped_seurat@meta.data

# Set up metadata as desired for aggregation and DE analysis
refmapped_seurat <- SetIdent(refmapped_seurat, value = factor(gsub("\\_"," ", refmapped_seurat@meta.data$predicted.celltype.l2)))
metadata$cluster_id <- factor(refmapped_seurat@active.ident)
metadata <- metadata %>%
  dplyr::rename(sample_id = "Sample.ID",
                class = "batch") %>%
  mutate(class = factor(tolower(class))) %>%
  mutate(experiment_mode = factor(gsub(".*a\\_", "", sample_id))) %>%
  mutate(sample_id = factor(gsub("\\_.*", "", tolower(sample_id))))

# Create single cell experiment object
sce <- SingleCellExperiment(assays = list(counts = counts), 
                           colData = metadata)

#change colnames of coldata to more r friendy names
colnames(colData(sce)) <- gsub("\\.","_",  colnames(colData(sce)))

# Identify groups for aggregation of counts
groups <- colData(sce)[, c("cluster_id", "sample_id", "predicted_celltype_l2")] #predicted_celltype_l2 is lower level clustering
```

### acquiring necessary metrics for aggregation across cells in a sample

```{r, acquire metrics}
# Named vector of cluster names
kids <- purrr::set_names(levels(sce$cluster_id))
kids

# Total number of clusters
nk <- length(kids)
nk

# Named vector of sample names
sids <- purrr::set_names(levels(sce$sample_id))

# Total number of samples 
ns <- length(sids)
ns

# Generate sample level metadata

## Determine the number of cells per sample
table(sce$sample_id)
table(sce$experiment_mode)
## Turn named vector into a numeric vector of number of cells per sample
n_cells <- as.numeric(table(sce$sample_id))

## Determine how to reorder the samples (rows) of the metadata to match the order of sample names in sids vector
m <- match(sids, sce$sample_id)

## Create the sample level metadata by combining the reordered metadata with the number of cells corresponding to each sample.
ei <- data.frame(colData(sce)[m, ], 
                  n_cells, row.names = NULL) %>% 
                select(-"cluster_id")
ei %>%
  dplyr::select(class, sample_id, n_cells)
```

### Remove low quality genes 

(re-read vignette to important points)

```{r, remove low quality using scater}
#skip because can't install scater
# Perform QC if not already performed
dim(sce)

# Calculate quality control (QC) metrics
#sce <- calculateQCMetrics(sce)

# Get cells w/ few/many detected genes
#sce$is_outlier <- isOutlier(
#        metric = sce$total_features_by_counts,
#        nmads = 2, type = "both", log = TRUE)

# Remove outlier cells
#sce <- sce[, !sce$is_outlier]
```

```{r, remove lowly expressed genes}
## Remove lowly expressed genes which have less than 10 cells with any counts
dim(sce)
sce <- sce[rowSums(counts(sce) > 1) >= 10, ]
dim(sce)
```

## Count aggregation to sample level

```{r, count aggregation to sample level}
# Aggregate the counts per sample_id and cluster_id

# Subset metadata to only include the cluster and sample IDs to aggregate across
groups <- colData(sce)[, c("cluster_id", "sample_id")]

# Aggregate across cluster-sample groups
pb <- aggregate.Matrix(t(counts(sce)), 
                       groupings = groups, fun = "sum") 

class(pb)

dim(pb)

pb[1:6, 1:6]
```

### Split data by cell type

```{r split data by cell type}
# Not every cluster is present in all samples; create a vector that represents how to split samples
splitf <- sapply(stringr::str_split(rownames(pb), 
                                    pattern = "_(?!.*_)",
                                    n = 2), 
                 `[`, 1)

# Turn into a list and split the list into components for each cluster and transform, so rows are genes and columns are samples and make rownames as the sample IDs
pb <- split.data.frame(pb, 
                       factor(splitf)) %>%
        lapply(function(u) 
                set_colnames(t(u), 
                             stringr::str_extract(rownames(u), "(?<=_)[:alnum:]+")))

class(pb)

# Explore the different components of list
str(pb)
```

### Check counts per sample for each cluster

```{r, check counts per sample for each cluster}
# Print out the table of cells in each cluster-sample group
options(width = 100)
table(sce$cluster_id, sce$sample_id)
colSums(table(sce$cluster_id, sce$sample_id))

cell_types_above_200 <- levels(sce$cluster_id)[colSums(t(table(sce$cluster_id, sce$sample_id))) > 200] #limit to cell types with enough cells
```

## Prepare sample-level metadata for DE anaylsis

```{r, prepare sample level metadata}
# Get sample names for each of the cell type clusters

# prep. data.frame for plotting
get_sample_ids <- function(x){
        pb[[x]] %>%
                colnames()
}

de_samples <- map(1:length(kids), get_sample_ids) %>%
        unlist()

# Get cluster IDs for each of the samples

samples_list <- map(1:length(kids), get_sample_ids)

get_cluster_ids <- function(x){
        rep(names(pb)[x], 
            each = length(samples_list[[x]]))
}

de_cluster_ids <- map(1:length(kids), get_cluster_ids) %>%
        unlist()

# Create a data frame with the sample IDs, cluster IDs and condition
gg_df <- data.frame(cluster_id = de_cluster_ids,
                    sample_id = de_samples)

gg_df <- left_join(gg_df, ei[, c("sample_id", "class")]) 


metadata <- gg_df %>%
        dplyr::select(cluster_id, sample_id, class) 
        
metadata 
```

### Subsetting dataset to cluster(s) of interest

```{r subset to clusters of interest}
# Generate vector of cluster IDs
clusters <- levels(factor(metadata$cluster_id))
clusters
```

Subset on celltype of interest and run DESeq2 as a loop

```{r subset on celltype of interest then run DESeq2 all in a loop}
res_tbl <- res <- rld_cor <- rld_mat <- rld <- dds <- counts <- cluster_counts <- cluster_metadata <- sapply(cell_types_above_200, function(x) NULL)
for(i in cell_types_above_200){
  cluster_metadata[[i]] <- metadata[which(metadata$cluster_id == i), ]
  #head(cluster_metadata[[i]] )
  # Assign the rownames of the metadata to be the sample IDs
  rownames(cluster_metadata[[i]]) <- cluster_metadata[[i]]$sample_id
  #head(cluster_metadata[[i]] )
  # Subset the counts to only the "Mono"
  counts[[i]] <- pb[[i]]
  cluster_counts[[i]] <- data.frame(counts[[i]][, which(colnames(counts[[i]]) %in% rownames(cluster_metadata[[i]]))])
  # Check that all of the row names of the metadata are the same and in the same order as the column names of the counts in order to use as input to DESeq2
  if(all(rownames(cluster_metadata[[i]]) == colnames(cluster_counts[[i]]))){
    print(paste0("all good for ", i, "!"))
  }
  dds[[i]] <- DESeqDataSetFromMatrix(cluster_counts[[i]], 
                              colData = cluster_metadata[[i]], 
                              design = ~ class)
  dds[[i]]  <- estimateSizeFactors(dds[[i]] )
  dds[[i]]  <- estimateDispersions(dds[[i]] )
  dds[[i]]  <- nbinomWaldTest(dds[[i]] )
  print(paste0("comparison for ", i, " is ", resultsNames(dds[[i]])[2]))
  print(paste0("showing plots for ", i))
  rld[[i]] <- rlog(dds[[i]], blind=TRUE) #Transform counts for data visualization
  DESeq2::plotPCA(rld[[i]], intgroup = "class") # Plot PCA
  rld_mat[[i]] <- assay(rld[[i]]) # Extract the rlog matrix from the object and compute pairwise correlation values
  rld_cor[[i]] <- cor(rld_mat[[i]])
  pheatmap::pheatmap(rld_cor[[i]], annotation = cluster_metadata[[i]][, c("class"), drop=F]) # Plot heatmap
  dds[[i]] <- DESeq(dds[[i]]) # Run DESeq2 differential expression analysis
  plotDispEsts(dds[[i]]) # Plot dispersion estimates (we expect our dispersions to decrease with increasing mean and follow the line of best fit)
  res[[i]] <- results(dds[[i]],
                 contrast = c("class", levels(cluster_metadata[[i]] $class)[2], levels(cluster_metadata[[i]]$class)[1]),
                 alpha = 0.05)
  res[[i]] <- lfcShrink(dds[[i]],
                        coef =  2,
                        type = "apeglm",
                        res=res[[i]])
  res_tbl[[i]] <- res[[i]] %>%
    data.frame() %>%
    rownames_to_column(var="gene") %>%
    as_tibble() %>%
    arrange(pvalue) %>%
    mutate(comparison = resultsNames(dds[[i]])[2])
}

for(i in cell_types_above_200){
  res_tbl[[i]] <- res_tbl[[i]] %>%
    mutate(ranking_metric = -log10(pvalue)*sign(log2FoldChange)) %>%
    arrange(desc(ranking_metric)) %>%
    drop_na()
}


```

### Bind all data frames into single dataframe with cell_type column

```{r, de tables}
# Turn the results object into a tibble for use with tidyverse functions
res_tbl_bound <- bind_rows(res_tbl, .id = "cell_type")

# Write all results to file
write.csv(res_tbl_bound,
          paste0("/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/Kalifabougou Sterile Immunity Manuscript/R21-sc-studies/DE_analysis_scrnaseq/results/pseudobulk_DESeq2_lower_level_celltypes_all_DEG_tables.csv"),
          quote = FALSE, 
          row.names = FALSE)
```
